Something about Performance, 2D Component of NEMO. (TBC)

### 2016-08-19 nemolite2d_perf.tex
nemolite-perf: added some conclusions!

The introduction of these directives proved to be simpler than
producing the OpenMP version of NEMOLite2D {\bf - Jeremy to clarify
  whether this is true}.  Consequently we only show the performance of
{\bf xxx} versions of the code in Figure~\ref{FIG_gpu_opt_stages}.

@@

We have investigated the application of the \psykal\ separation of
concerns approach to the domain of shared-memory parallel,
finite-difference ocean models. This approach enables the
computational science (performance) related aspects of a computer
model to be kept separate from the natural (oceanographic) science
aspects.

We have used a new and un-optimised, two-dimension model extracted
from the NEMO ocean model for this work. As a consequence of this, the
introduction of the \psykal\ separation of concerns followed by
suitable transformations of the PSy Layer is found to improve
performance. This is in contrast to our previous
experience~\cite{shallow_psykal} with tackling the Shallow code which
has had been optimised over the course of several years. In that case
we were able to recover\ldots

The application of code transformations to the middle/PSy layer is key
to the performance of the \psykal\ version of a code.

>>>>

The introduction of these directives proved to be simpler than
producing the OpenMP version of NEMOLite2D {\bf - Jeremy to clarify
  whether this is true}.  Consequently we only show the performance of
{\bf xxx} versions of the code in
Figure~\ref{FIG_gpu_opt_stages}. {\bf discuss performance drop in
  OpenACC version c.f. CUDA.}

@@

We have investigated the application of the \psykal\ separation of
concerns approach to the domain of shared-memory parallel,
finite-difference shallow-water models. This approach enables the
computational science (performance) related aspects of a computer
model to be kept separate from the natural (oceanographic) science
aspects.

We have used a new and un-optimised, two-dimension model extracted
from the NEMO ocean model for this work. As a consequence of this, the
introduction of the \psykal\ separation of concerns followed by
suitable transformations of the PSy Layer is actually found to improve
performance. This is in contrast to our previous
experience~\cite{shallow_psykal} with tackling the Shallow code which
has had been optimised over the course of several years. In that case
we were able to recover more than 90\% of the performance of the
original for the Gnu, Intel, Cray and IBM compilers on both Intel and
IBM Power CPUs.

The application of code transformations to the middle/PSy layer is key
to the performance of the \psykal\ version of a code. For both
NEMLite2D and Shallow we have found that for serial performance, the
most important transformation is that of in-lining the kernel source
at the call site. However, for the more complex NEMOLite2D code the
Cray compiler also had to be coerced into performing SIMD
vectorisation through the use of souce-code directives.

In this work we have also demonstrated the introduction of parallelism
into the PSy layer with both OpenMP and OpenACC directives. In both
cases we were able to leave the natural-science parts of the code
unchanged. For OpenMP we achieved a parallel efficiency of $\sim70\%$
on 12 threads by enclosing the body of the (single) PSy layer routine
within a single PARALLEL region. Removal of unecessary synchronisation
points through use of the NOWAIT clause boosted 12-thread performance
by approximately 10\% with the Gnu and Cray compilers. The
\psykal\ re-structuring of this (admittedly small) code was not found
to pose any problems for the introduction of performant OpenMP.

{\bf Some words about how good the OpenACC, \psykal\ version is compared
to the CUDA version.}

Having demonstrated that the \psykal\ separation of concerns may be
applied to finite-difference codes with minimal performance loss, our
next step will be to consider the automatic generation of the PSy
layer. In future work we will analyse the performance of a
domain-specific compiler that performs this task. This compiler
(which we have named `PSyclone') is currently under development.
